---
title: "LEAP 2017 metacommunities -- report"
author: "Vincent Fugère, January 2020"
output: pdf_document
urlcolor: blue
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include=FALSE)
```

```{r}
rm(list=ls())
```

```{r, load packages}
library(tidyverse)
library(readxl)
library(magrittr)
library(scales)
library(chron)

library(brms)
library(vegan)
library(mgcv)
library(itsadug)
library(ape)

source('/Users/vincentfugere/Google Drive/Recherche/PhD/R/functions/utils.R')
```

### Effects of pH on zooplankton ###

Lets first look at this in the Thibodeau dataset, quantifying for each bag: 1) total zooplankton density, 2) species richness, 3) alpha diversity (Hill number, or exponent of Shannon index), and 4) community composition. I could not fit an ordination to this data because there are only 9 mesocosms with 7 species, and treatment effects are very strong (such that many points sit on top of each other in multivariate space). By simply looking at the community matrix, I noticed that: 1) Nauplii disappear completely in the press treatment but increase in abundance to > 95% relative abundance in the pulse treatment; 2) the cladocerans *Bosmina* and *Ceriodaphnia* disappear completely in both acid treatments; 3) in the press treatment, each bag is dominated (>80% relative abundance) by a different cladoceran, one of: _Chydorus_, _Daphnia_, or _Diaphanosoma_. Density and diversity are also affected by the treatments. These results are shown in Figure 1.

```{r}
source(here::here('./thibodeau_zoo.R'))
```

```{r, include=TRUE, fig.width=5,fig.height=6,fig.cap="Impacts of acidification on zooplankton communities in in-lake mesocosms"}
par(mfrow=c(3,2),cex=0.7,mar=c(4,4,0.5,0.5))
boxplot(density~treatment,thibodeau,ylab='total density')
boxplot(richness~treatment,thibodeau,ylab='richness')
boxplot(hill~treatment,thibodeau,ylab=expression(e^Shannon))
boxplot(Nauplii~treatment,thibodeau,ylab='% nauplii')
boxplot(bosm_cerio~treatment,thibodeau,ylab='% bosm+cerio')
boxplot(chyd_daph_diaphan~treatment,thibodeau,ylab='% chyr+daph+diaph')
par(mfrow=c(1,1))
```

```{r}
source(here::here('./LEAP_data_format.R'))
```

Lets now do a similar analysis with the LEAP data, only including isolated ponds from metacommunities without dispersal. This shows the effect of pH in the absence of dispersal. Again, we look at density, diversity, and composition.


